%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% I, the copyright holder of this work, release this work into the
%% public domain. This applies worldwide. In some countries this may
%% not be legally possible; if so: I grant anyone the right to use
%% this work for any purpose, without any conditions, unless such
%% conditions are required by law.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[
  digital, %% The `digital` option enables the default options for the
           %% digital version of a document. Replace with `printed`
           %% to enable the default options for the printed version
           %% of a document.
%%  color,   %% Uncomment these lines (by removing the %% at the
%%           %% beginning) to use color in the digital version of your
%%           %% document
  table,   %% The `table` option causes the coloring of tables.
           %% Replace with `notable` to restore plain LaTeX tables.
  twoside, %% The `twoside` option enables double-sided typesetting.
           %% Use at least 120 g/m² paper to prevent show-through.
           %% Replace with `oneside` to use one-sided typesetting;
           %% use only if you don’t have access to a double-sided
           %% printer, or if one-sided typesetting is a formal
           %% requirement at your faculty.
  nolof,     %% The `lof` option prints the List of Figures. Replace
           %% with `nolof` to hide the List of Figures.
  nolot,     %% The `lot` option prints the List of Tables. Replace
           %% with `nolot` to hide the List of Tables.
  %% More options are listed in the user guide at
  %% <http://mirrors.ctan.org/macros/latex/contrib/fithesis/guide/mu/fi.pdf>.
  draft=false,
  final,
]{fithesis3}
%% The following section sets up the locales used in the thesis.
\usepackage[resetfonts]{cmap} %% We need to load the T2A font encoding
\usepackage[T1,T2A]{fontenc}  %% to use the Cyrillic fonts with Russian texts.
\usepackage[
  main=english, %% By using `czech` or `slovak` as the main locale
                %% instead of `english`, you can typeset the thesis
                %% in either Czech or Slovak, respectively.
  english, german, russian, czech, slovak %% The additional keys allow
]{babel}        %% foreign texts to be typeset as follows:
%%
%%   \begin{otherlanguage}{german}  ... \end{otherlanguage}
%%   \begin{otherlanguage}{russian} ... \end{otherlanguage}
%%   \begin{otherlanguage}{czech}   ... \end{otherlanguage}
%%   \begin{otherlanguage}{slovak}  ... \end{otherlanguage}
%%
%% For non-Latin scripts, it may be necessary to load additional
%% fonts:
\usepackage{paratype}
% my packages
\usepackage{framed,color}
\usepackage{textcmds}
\def\textrussian#1{{\usefont{T2A}{PTSerif-TLF}{m}{rm}#1}}
%%
%% The following section sets up the metadata of the thesis.
\thesissetup{
    date        = \the\year/\the\month/\the\day,
    university  = mu,
    faculty     = fi,
    type        = bc,
    author      = Roman Solař,
    gender      = m,
    advisor     = {doc. RNDr. Pavel Matula, Ph.D.},
    title       = {Generating of Mathematical Problems for the Course PV189},
    TeXtitle    = {Generating of Mathematical Problems for the Course PV189},
    keywords    = {PV189, mathematics for computer graphics, SymPy, ROPOT question set, mathematical problem generation},
    TeXkeywords = {PV189, mathematics for computer graphics, SymPy, ROPOT question set, mathematical problem generation},
    abstract    = {%
      This thesis aims to generate question sets for the course PV189 while also providing a functional interface for future amendments and additions. The output must be compatible with IS ROPOTs, and it should be able to set a tolerance for non-integer answers. It should be possible to solve all the generated questions by hand, possibly with the aid of a calculator.  
The tasks generally revise basic concepts from the lectures and help build geometric intuition about them.  For many questions, symbolic computation has proven useful. In this case, the full functionality is implemented using Python 3; the symbolic computation is accomplished via the open-source Python library SymPy. 
    },
    thanks      = {%
      I take this opportunity to express gratitude to my supervisor doc. RNDr. Pavel Matula, Ph.D. and appreciate him for not losing his patience with me. Additionally, I would like to say a special thanks to Eva Drštková for going the extra mile to make this thesis possible. 
    },
    bib         = example.bib,
    %% Uncomment the following line (by removing the %% at the
    %% beginning) and replace `assignment.pdf` with the filename
    %% of your scanned thesis assignment.
%%    assignment         = assignment.pdf,
}
\usepackage{makeidx}      %% The `makeidx` package contains
\makeindex                %% helper commands for index typesetting.
%% These additional packages are used within the document:
\usepackage{paralist} %% Compact list environments
\usepackage{amsmath}  %% Mathematics
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{url}      %% Hyperlinks
\usepackage{markdown} %% Lightweight markup
\usepackage{tabularx} %% Tables
\usepackage{tabu}
\usepackage{booktabs}
\usepackage[draft=false]{hyperref}
\usepackage{listings} %% Source code highlighting
\lstset{
  basicstyle      = \ttfamily,
  identifierstyle = \color{black},
  keywordstyle    = \color{blue},
  keywordstyle    = {[2]\color{cyan}},
  keywordstyle    = {[3]\color{olive}},
  stringstyle     = \color{teal},
  commentstyle    = \itshape\color{magenta},
  breaklines      = true,
}
\usepackage{floatrow} %% Putting captions above tables
\floatsetup[table]{capposition=top}
\thesissetup{assignment={}}
\begin{document}

\chapter{Introduction}

\input{introduction}


\chapter{Tools}

Throughout the implementation of the project, I had to familiarize myself with new tools in order to accomplish or facilitate different parts of the project. This includes well-known libraries \verb|numpy| and \verb|scipy| with plentiful documentation, and for this reason I will not expand on them further. Aside from these, I have frequently made use of the library for symbolic computation \verb|SymPy|, the basics of which are explained further in this section. 

As the purpose of this project is to create question sets compatible with the university information system IS, it is necessary to explain the basic syntax for creating question sets. 

\section{SymPy}

\input{sympy}

\section{ROPOTs}

\input{ropots}

\chapter{Using the project}

\shorthandoff{-}
\begin{markdown*}{%
  hybrid,
  definitionLists,
  footnotes,
  inlineFootnotes,
  hashEnumerators,
  fencedCode,
  citations,
  citationNbsps,
  pipeTables,
  tableCaptions,
}

There are two basic scenarios in which you would need to work with the project. The first one is when you use it to generate questions, e.g., when you want to update a set of questions because the students have memorized the answers already. In order to do it, you need to know the structure of the project. In the second scenario, you create or alter a task. In that case, you need to understand the structure of the project as well as implementation details. You should know where to put the new task, what you need to fill it with, and how to do it in accordance with the project best practices. 

\section{Generating question sets}

*Note that this section does not aim to dive deep into details; instead, it provides a quick overview of the standard workflow. It is handy if you already have a basic understanding of the provided methods.*

You want to generate a new question set for the desired task. You need to know what task this is and the lecture it is in. 

1. In the project tree, navigate to the folder corresponding to the lecture covering the task you want to work with.
    
2. Find the python file corresponding to the task inside the folder. The filename should match the task name.
    
3. Execute the file. This will load all necessary dependencies and the class corresponding to the task. The class name is equal to the task name. 
    
4. Once the class is loaded, you can generate the questions calling its methods. It is recommended to call
``` Python
>>> TaskClassName.prepare_n_questions(n,  limits, tolerance, filename, points)
```
where you replace ```TaskClassName``` with the actual name of the class. 
    * argument **```n```** specifies the number of questions to be created
    * argument **```limits```** specifies the limits for the generated questions
    * argument **```tolerance```** specifies the tolerance for the numerical answers (typically 0.0001)
    * argument **```filename```** specifies the path to a file to store the generated questions
    * argument **```points```** specifies the number of points for each answer field (as a list of points respecting the order of the answer fields)
    
#. You can find the generated questions inside the file **```filename```**. Copy and paste them directly inside a question set in IS.

\subsection{Quick example}

Here is a quick Python console example for generating a question set for a concrete task (`ScalarProductFromAngle`).  

```Python
>>> ScalarProductFromAngle.prepare_n_questions(20, [-5, 5], 0.0001,                         "scalar_product_from_angle_questions.txt")
```

This command generates 20 questions inside the file \newline `scalar_product_from_angle_questions.txt`. The tolerance for the answers is 0.0001 and the limits are `[-5, 5]`, the points are not specified so the task default will be used. 

\section{Testing task implementation}

You want to make sure the implemented task produces valid instances. 

1. In the project tree, navigate to the folder corresponding to the lecture covering the task you want to work with.
    
2. Find the python file corresponding to the task inside the folder. The filename should match the task name.
    
3. Execute the file. This will load all necessary dependencies and the class corresponding to the task. The class name is equal to the task name. 
    
4. Once the class is loaded, you can generate a list of instances calling its methods. Save the result inside a variable to use it later.
``` Python
>>> instances = TaskClassName.get_n_unique_instances(n, limits, max_rounds=100)
```
where you replace ```TaskClassName``` with the actual name of the class. 
    * argument **```n```** specifies the number of different instances to be created
    * argument **```limits```** specifies the limits for the generated instances
    * argument **```max_rounds```** specifies the maximal number of unsuccessful attempts to retrieve a different instance in a row (the instances are typically randomly generated)
    * the method returns a list of `n` unique instances

5. Now, it is possible to test the validity of the generated instances. 
``` Python
>>> TaskClassName.test_instances(instances)
```
where you replace ```TaskClassName``` with the actual name of the class. 
    * argument **```limits```** takes a list of instances to be tested
    * the method prints out the number of invalid instances, and the concrete instances that failed the tests
    * the method returns `True` if all the instances are valid, `False` otherwise


\subsection{Quick example}

Here is a quick Python console example for testing a concrete task (`ScalarProductFromAngle`). 

```Python
>>> instances = ScalarProductFromAngle.get_n_unique_instances(100, [-5, 5])
>>> ScalarProductFromAngle.test_instances(instances)
0 errors in 100 instances
True
```


\chapter{Development guide}

You should read this chapter if you want to implement your own task or understand the very details of the project implementation. This is the main documentation of the project, but you can find more documentation inside the project files and the project \verb|README|. This might help, if something seems unclear. As a last resort, you can always check the project implementation.

\section{Creating a new task}

Before creating a new task, make sure you understand the difference between a question and a task. You should know what an instance is and how methods work with it. If this is not the case, I recommend starting out with \hyperref[sec:example]{the example}. 

\subsection{Updating the project tree}

#. Start by creating a **new file** for the task in the folder corresponding to the lecture
the task belongs to. (Folder name '*LX*' stands for '*Lecture X*'.) 
\definecolor{shadecolor}{rgb}{1,0.8,0.3}
\begin{shaded}{**Best practice:**}
    The strict rule is one task per one file.
\end{shaded}

#. Create a **new class** in the file to represent the task. The class must be a child of
the common parent class `Task` (located in the root of the project in `task.py`).

#. Choose a descriptive **name for the task**.
\definecolor{shadecolor}{rgb}{1,0.8,0.3}
\begin{shaded}{**Best practice:**}
    The name of the class should be in the `CapitalizedWords` style, the **name of the file** should be the same but in the `lower_case_with_underscores` style.
\end{shaded}

\subsection{Implementing abstract methods}

The listed abstract methods are obligatory to implement. If something is missing here,
you can check out the corresponding documentation in the class `Task`. 
It is recommended to implement the methods in the described order.

\subsubsection{Abstract method genfun} 

This is the core function of the task, it is responsible for **creating a valid
instance** of the problem. The instance of the problem is a term coined to describe
the concrete values of the variables of the problem and its solution.

The only parameter of the method, `limits`, can be used to **limit the randomly
generated values** of the problem variables. There are no restrictions on the
parameter --- it can be a list of numbers, a dictionary... However, in almost 
every case, it is a list of two numbers, where the first number represents the 
minimum value, and the second number the maximum value for a random generator.

\definecolor{shadecolor}{rgb}{1,0.8,0.3}
\begin{shaded}{**Best practice:**}
    Sometimes, it is necessary to leave out some numbers from the range (such as 0 or negative numbers).
In this case, the method should be robust enough to ignore undesired values within the range.
\end{shaded}

The function **returns a valid instance** of the problem, typically randomly generated.
However, the instance **must be hashable** (it has a `__hash__()` method) and so all lists inside an instance should be replaced by tuples etc.


\subsubsection{Abstract method testfun} 

This method takes an **instance** as a parameter and tests if it is **valid**. Since this method is not generally used when generating questions, the developer has a much more free hand implementing this method.

In general, the tests should sanity check the values in the instance (e.g., when asked to compute the division of two numbers `a` and `b`, we check that `b` is not equal to 0), and check that the result is compatible with the other values. Naturally, checking the latter should be done using different techniques (typically different libraries) than inside `genfun` or `solve_instance`.

The method should ideally **never fail** or raise an exception, instead it **returns**
`True` if the instance is valid and `False` otherwise.

Sometimes, it is impossible to meet the aforementioned criteria~---~usually because there is no other widespread trustworthy library compatible with the format of the instance. In that case, the test can either be partial or completely skipped. In case you want to skip implementing the method `testfun`, make it always return `True` and provide a comment explaining why you have decided to do so.

\subsubsection[Abstract method create_question]{Abstract method `create_question`} 

This method takes a valid instance (returned by `genfun`) and turns it into a question,
which can be passed into a ROPOT in IS. 

In order to be able to do this, it must know the **tolerance** for the answer (typically $0.0001$),
and the **points** given for each answer (there might be multiple input fields).

It **returns a string** which contains the question for the ROPOT.

\subsection{Finishing the implementation}

*This section only applies to you if you intend to make your changes to the project public.*

Because in the final version, there should be no easy guideline
for students to solve the given questions, we need to "hide" the
result-computing logic of the `genfun` method inside a new static method 
`solve_fun`. In the published version, the body of this new method will be 
replaced with the keyword `pass`.

\definecolor{shadecolor}{rgb}{1,0.8,0.3}
\begin{shaded}{**Best practice:**}
Before you finish the implementation of the new task, you should document it. Each task class documentation should explain
\begin{itemize}
    \item the task itself (what problem the student is supposed to solve),
    \item what an instance of the task is composed of and 
    \item the role of the argument `limits`.
\end{itemize}
\end{shaded}

\section{Example}
\label{sec:example}

*In this section, I will use a very primitive example to demonstrate how the 
development of a new task should ideally work.*

The example problem is to find the following number in the arithmetic sequence, 
given by its first two terms.

### Creating a new task

The students should have mastered this topic already, and we do not need to cover 
it in any lecture. Therefore, we will create a new folder `L0` and place a new
file `arithmetic_sequence.py` in it. Inside the file, I will create the following template.

```python
import random as rnd

from task import Task

class ArithmeticSequence(Task):

    @classmethod
    def create_question(cls, instance, tol, points):
        pass
    
    @classmethod
    def genfun(cls, limits):
        pass
    
    @classmethod
    def testfun(cls, instance):
        pass
```

### Implementing abstract methods

First, I will implement the `genfun` method. The instance for my problem
can be represented by the first two terms of the sequence, and the third
term, which is the answer. I choose to use the limits to represent the lowest
value, and the largest value for the first two terms of the sequence.

```python
def genfun(cls, limits):
    min_val, max_val = limits
    vals = list(range(min_val, max_val + 1))
    a1, a2 = rnd.choice(vals), rnd.choice(vals)

    difference = a2 - a1
    a3 = a2 + difference
    return a1, a2, a3
```

In the last step of the method we return the instance (which is composed of the
first three terms in this case).

Now, we implement `testfun` to make sure our instances are valid.

```python
def testfun(cls, instance):
    a1, a2, a3 = instance  # "unpack" the instance
    return a2 - a1 == a3 - a2
```

Finally, we implement `create_question` to transform instances
into questions.

```python
def create_question(cls, instance, tol, points=[1]):
    a1, a2, a3 = instance
    return f"""
    Find the third term of an arithmetic sequence, if the first terms 
    are: {a1}, {a2}. 
    
    <m>a_3 = </m> :n
    :n={Task.answer(a3, 4, tol)} {points[0]} ok
    """
```

This will create a question for a given instance, if we also set the
tolerance `tol` for the answer (in this case it could easily be `0`, typically `0.0001`).
In this case, the student will receive 1 point for a correct answer (within the given tolerance).

Now, I will separate the code generating the answer from the `genfun` method and
"hide" it inside the `solve_fun` method, so it can be easily removed later in the final version.

In the end, the file `arithmetic_sequence.py` will look like this:

```python
import random as rnd

from task import Task

class ArithmeticSequence(Task):

    @classmethod
    def create_question(cls, instance, tol, points=[1]):
        a1, a2, a3 = instance
        return f"""
        Find the third term of an arithmetic sequence, if the first terms 
        are: {a1}, {a2}. 
        
        <m>a_3 = </m> :n
        :n={Task.answer(a3, 4, tol)} {points[0]} ok
        """
    
    @staticmethod
    def solve_instance(a1, a2):
        difference = a2 - a1
        a3 = a2 + difference
        return a3
    
    @classmethod
    def genfun(cls, limits):
        min_val, max_val = limits
        vals = list(range(min_val, max_val + 1))
        a1, a2 = rnd.choice(vals), rnd.choice(vals)
    
        a3 = cls.solve_instance(a1, a2)
        return a1, a2, a3
    
    @classmethod
    def testfun(cls, instance):
        a1, a2, a3 = instance  # "unpack" the instance
        return a2 - a1 == a3 - a2
```

### Using the new class

From the Python CLI, I will load the class `ArithmeticSequence` and test it.
```python
>>> ArithmeticSequence.test(1000)
0 errors in 1000 instances
```

Because there seem to be no errors, I proceed to generate 20 questions with it.
Before that, I need to create a new folder `questions` inside the `L0` folder.
```python
>>> ArithmeticSequence.questions()
```

I can check the outcome inside the file `questions/ArithmeticSequence.txt`, which
can be directly passed into a ROPOT.

\end{markdown*}
\shorthandon{-}








\chapter{Glossary}

All these keywords are explained in the text of the thesis when they are first introduced. However, it might be difficult to look for them if you do not read the text sequentially.


\paragraph{instance}

A collection of data that uniquely represents a given question, including the answer to it. E.g., for the question ``\textit{Compute the sum $1 + 3$}'', the instance could be \textit{1, 3, 4.}

\paragraph{IS}

The information system for MUNI, available at \href{https://is.muni.cz/}{https://is.muni.cz/}.

\paragraph{question}

A human-readable text containing the problem formulation and all the necessary data to solve it. It has a unique answer. E.g., the text ``\textit{Compute the sum $1 + 3$}.'' is a question.


\paragraph{question set}

This is usually a file containing one or more questions. An example can be found inside the root of the project: \newline \verb|sample_question_set.txt|.

\paragraph{ROPOT}

\textbf{R}evision \textbf{O}pinion \textbf{PO}ll and \textbf{T}esting is an online application in IS capable of assigning problems to students and evaluating their answers. It can parse question sets.


\paragraph{task}

The implementation for generation of a concrete problem. E.g., the class \verb|ImplicitLineFromTwoPoints|.


\appendix

\input{implemented_tasks}


\chapter{Contents}


\end{document}
